---
title: "Fenton z-score ~ Metabolomics"
output: 
  html_document:
    df_print: paged
    number_sections: true
date: "`r format(Sys.Date(),'%B %d, %Y')`"
---
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(data.table)
library(readxl)
library(corrplot)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(psych)
library(caret)
library(ggfortify)
library(doParallel)
library(tidyverse)
library(bayesSurv)
library(GIGrvg)
library(statmod)
library(MCMCpack)
library(reshape2)
library(latex2exp)
library(parallel)
library(infinitefactor)
library(viridis)
library(robustbase)
library(BWQS)
library(DT)
library(table1)
library(Rtsne)
library(VIM)
library(mice)
library(skimr)
library(ggsci)
library(apcluster)
library(MASS)
library(gclus)
library(qgraph)
```

# Dataset

## Variable description

Outcome: 

+ **bw_z_Fenton**: birth weight for gestational age, Fenton z-score

Auxiliary variables:

+ **hhid**: household ID (50xxx: Boston, 70xxx: NY)
+ **iid**: individual ID
+ **ch_death**: child death at some point: 0=No, 1=Yes
+ **bw**: child's birth weight (g)
+ **educ_low**: 1 = high school or less, 0 = above high school degree
+ **mbmi_prepreg**: maternal pre-pregnant BMI (kg/m^2^)
+ **parity**: number of previous births, 0=0, 1=1, 2= >1
+ **smk_preg**: maternal smoking during pregnancy: 0=no, 1=yes
+ **gestage**: gestational age at delivery (weeks)

Covariates: 

+ **childsex**: child sex --> 0=female (reference), 1=male
+ **momrace2**: 0=white (reference), 1=black, 2=hispanic, 3=other
+ **educ**: 1= <12th grade (reference), 2= hs degree or ged, 3= some college or associates, 4= 4 years of college, 5= graduate degree
+ **age_birth**: maternal age at birth

Metabolites:

```{r metabolites_list, warning=FALSE, message=FALSE, echo=F}
# Data import
met = fread("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/3_metabolites.csv")
metabolites_list = colnames(met)[-c(1,2)]
met_description = fread(file = "J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/1_metabolite-details.csv")
met_description = met_description[comp_id %in% metabolites_list,.(comp_id, biochemical, super_pathway, 
                                                sub_pathway, platform, chemical_id,
                                                kegg,group_hmdb)]

met_matrix_description = as.matrix(met_description[,-1,with=F])
colnames(met_matrix_description) = c("Biochemical","Super Path",
                                     "Sub Path","Platform","Chem ID",
                                     "KEGG","HMDB")
rownames(met_matrix_description) = met_description$comp_id  
datatable(met_matrix_description, filter = 'top', options = list(
  pageLength = 5, autoWidth = TRUE,lengthChange = FALSE))
```

## Missing Data and Imputation

Summary of missing data on the covariates of interest:

```{r import, echo=F, warning=FALSE, message=FALSE, fig.align='center'}
dt = fread("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/EC20190917.csv")
dt = dt[,1:39,with=F]
dt[,iid := ifelse(iid<10,iid*10,iid)]
dt[,studyID_2 := paste0(hhid,"-",iid)]

met = fread("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/3_metabolites.csv")

# Merge the two dataset
DT = merge(dt, met, by="studyID_2")

# Select only the components that we want
metabolites_list = colnames(met)[-c(1,2)]
covariates = c("childsex","ch_death","bw","gestage","age_birth","momrace2",
               "educ","educ_low","mbmi_prepreg","parity","smk_preg","gestage_serum")
outcome = c("bw_z_Fenton")

# Exclude the missing outcome and the children death after birth
DT = DT[!is.na(bw_z_Fenton),c("hhid","iid",outcome,covariates,metabolites_list),with=F]
DT = DT[ch_death != 1]

# Analize missing data for the important covariates only
x = DT[,c(covariates[c(1,5,6,7,9,12)]),with=F]
x = DT[,c(covariates,outcome),with=F]

aggr(x, col=c('navyblue','yellow'),
     numbers=TRUE, sortVars=TRUE, combined = FALSE,
     labels=names(x), cex.axis=.7,
     gap=3, ylab=c("Missing data","Pattern"))


# Imputation with MICE package
imputed_dataset = mice(x)
x = complete(imputed_dataset,3)

# Fill the NA value, now we have all 410 complete observation
DT$childsex = x$childsex
DT$age_birth = x$age_birth
DT$momrace2 = x$momrace2
DT$educ = x$educ
DT$mbmi_prepreg = x$mbmi_prepreg
DT$gestage_serum = x$gestage_serum
 
# Create dummy variable for race
DT[,':='(white = ifelse(momrace2 == 0, 1, 0),
         black = ifelse(momrace2 == 1, 1, 0),
         hispanic = ifelse(momrace2 == 2, 1, 0),
         other = ifelse(momrace2 == 3, 1, 0))]

DT_unscaled = copy(DT)
# Scale and center metabolites
cols = metabolites_list
DT[, (cols) := lapply(.SD, scale, center = TRUE, scale = TRUE), .SDcols = cols]
DT[, bw_z_Fenton := scale(bw_z_Fenton,center = TRUE, scale = FALSE)]
```

## Summary statistics

```{r plots, echo=F, warning=FALSE, message=FALSE, fig.align='center'}
DT_plot = DT[,c(outcome,covariates),with=F]
DT_plot[,`Child Sex` := ifelse(childsex==0,"female","male")]
DT_plot[educ == 1,`Maternal Education` := "<12th grade"]
DT_plot[educ == 2,`Maternal Education` := "hs degree or ged"]
DT_plot[educ == 3,`Maternal Education` := "some college or associates"]
DT_plot[educ == 4,`Maternal Education` := "4 years of college"]
DT_plot[educ == 5,`Maternal Education` := "graduate degree"]
DT_plot[momrace2 == 0, `Maternal Race` := "white"]
DT_plot[momrace2 == 1, `Maternal Race` := "black"]
DT_plot[momrace2 == 2, `Maternal Race` := "hispanic"]
DT_plot[momrace2 == 3, `Maternal Race` := "other"]
DT_plot[, `BW Fenton z-score` := bw_z_Fenton]
DT_plot[, `Maternal Age` := age_birth]
table1(~ `BW Fenton z-score` + `Maternal Age` + `Child Sex` + `Maternal Education` + `Maternal Race`,
       data=DT_plot, render.continuous=c(.="Mean", "Quantiles"="Median [Q1, Q3]"), transpose = F)
list_plot = c(outcome,covariates)

# Plot putcome distribution 
plot(density(DT$bw_z_Fenton), xlab="",ylab="",main="Fenton z-score distribution")
qqnorm(DT$bw_z_Fenton, main = "Normal Q-Q Plot for Fenton z-score")
qqline(DT$bw_z_Fenton,datax = TRUE, lty=2)
ks.test(DT$bw_z_Fenton,"pnorm")
```

## Correlation

```{r corr, echo=F, warning=FALSE, message=FALSE, fig.align='center'}
cat("Pearson correlation matrix ordered by FPC (first principal component)")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F]), diag = FALSE, order = "FPC",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")

cat("Pearson correlation matrix ordered by AOE (angular order of the eigenvectors)")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")

cat("Pearson correlation matrix ordered by hclust")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F]), diag = FALSE, order = "hclust",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")

cat("Spearman correlation matrix ordered by FPC (first principal component)")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F], method = "spearman"), diag = FALSE, order = "FPC",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")

cat("Spearman correlation matrix ordered by AOE (angular order of the eigenvectors)")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F], method = "spearman"), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")

cat("Spearman correlation matrix ordered by hclust")
corrplot(cor(met[,-c(1,2,NCOL(met)),with=F], method = "spearman"), diag = FALSE, order = "hclust",
         tl.pos = "td", tl.cex = 0.01, method = "color", type = "upper", tl.col = "white")
```

# FIN Analysis

## Number of factor

```{r fin,echo=F, warning=FALSE, message=FALSE, fig.align='center'}
# Detect number of factor using PCA
number_component = 25
pc_met = prcomp(DT[,metabolites_list,with=F], 
                scale = T, center = T, retx = T)
dt_plot = data.frame(x = 1:NCOL(pc_met$rotation),y = summary(pc_met)$importance[2,])
tk = ggplot(data = dt_plot, aes(x=x,y=y)) +
  geom_point(aes(x=x,y=y)) + 
  geom_line(aes(x=x,y=y)) + 
  labs(title = "PCA Metabolites",
       subtitle = paste0("We consider only first ",number_component," components (",
                         round(100*summary(pc_met)$importance[3,number_component],0),
                         "% variance explained)"),
       caption = "Data source: Metabolomics") +
  geom_vline(xintercept = 20, linetype = "dashed", colour = "red") + 
  xlab('Components') + 
  ylab('% Var explained') +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 25, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size=15),
        plot.caption = element_text(hjust = 1),
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size = 10)) 
print(tk)
dt_plot$cum_var = cumsum(dt_plot$y)
colnames(dt_plot) = c("PCA_ID","Variance","Cumulative Variance")
```

## Bayesian Factor Analysis

```{r analysis,echo=F, warning=FALSE, message=FALSE, fig.align='center'}
# Run FIN analysis
y = as.matrix(DT$bw_z_Fenton,NROW(DT$bw_z_Fenton),1)
X = as.matrix(DT[,c(metabolites_list,covariates[c(5,7,9,12)],
                    "white","black","other","hispanic"),with=F]) 
nrun = 10000
burn = 9000
thin = 1
k = 25
# intdl = interactionDL(y = y, X = X, nrun = nrun, burn = burn, thin = thin, k = k)
# saveRDS(intdl,"J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/FIN_fenton_bw_25_scaled_bmi.rds")
intdl = readRDS("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/FIN_fenton_bw_25_scaled_bmi.rds")

# Rotate 
Rotout = jointRot(intdl$lambdaSamps, intdl$etaSamps)
mLambda = lmean(Rotout$lambda)
rownames(mLambda) = colnames(X)

# Create an array with all the simulation 
aLambda = array(NA,dim = c(1118,25,1000))
for(i in 1:1000){
  aLambda[,,i] = Rotout$lambda[[i]]
}

# Calculate quantile for each component in the factor
aLambdaq1 = apply(aLambda, c(1,2), quantile, c(0.025))
aLambdaq2 = apply(aLambda, c(1,2), quantile, c(0.975))


mLambda1 = as.data.table(mLambda)
setnames(mLambda1,names(mLambda1),paste0("factor_",1:25))
mLambda1[,names := colnames(X)]

aLambdaq1 = as.data.table(aLambdaq1)
setnames(aLambdaq1,names(aLambdaq1),paste0("factor_",1:25))
aLambdaq1[,names := colnames(X)]

aLambdaq2 = as.data.table(aLambdaq2)
setnames(aLambdaq2,names(aLambdaq2),paste0("factor_",1:25))
aLambdaq2[,names := colnames(X)]

mLambda1 = melt(mLambda1)
aLambdaq1 = melt(aLambdaq1)
aLambdaq2 = melt(aLambdaq2)

mLambda1 = as.data.table(mLambda1)
aLambdaq1 = as.data.table(aLambdaq1)
aLambdaq2 = as.data.table(aLambdaq2)

aLambdaq1 = aLambdaq1[,.(names, variable=as.character(variable), value)]
mLambda1 = mLambda1[,.(names, variable=as.character(variable), value)]
aLambdaq2 = aLambdaq2[,.(names, variable=as.character(variable), value)]

# Table of the loadings  
data = Reduce(function(x, y) merge(x, y, 
    by=c("names", "variable")), list(aLambdaq1,mLambda1,aLambdaq2))
data[,check := ifelse(sign(value.x) == sign(value),1,0)]
setnames(data,names(data),c("Metabolite","Factor","q025","mean","q975","sig"))
data[order(sig,decreasing = T)]

# --- Plot --- #
# number of variables to include in the plot
# take the variables with the n_var largest loadings(in absolute value)
mLambda_met = mLambda[-which((rownames(mLambda) %in% c("age_birth","educ","mbmi_prepreg",
                                                       "gestage_serum","white",
                                                       "black","other","hispanic"))),]
n_var = 1110
max_load = apply(mLambda_met %>% abs(), 1, max)
ind = max_load %>% order()
mLambda_plot = mLambda[ind,]
mLambda_plot = mLambda_plot[1:n_var,]

# rearrage the rows of Lambda for interpretability
ind_max = apply(mLambda_plot %>% abs(), 1, function(x) which(x == max(x))) %>% order()
mat = mLambda_plot[ind_max,]
mat = apply(mat, 2, rev)
longmat = as.data.table(melt(mat))
Var1 = Var2 = value = NULL
p = ggplot(longmat, aes(x = Var2, y = Var1)) + geom_tile(aes(fill = value)) +
  scale_fill_gradient2(low = "#3d52bf", high = "#33961b", 
            mid = "white") + 
  scale_x_discrete(limits = paste(1:25)) +
  coord_cartesian(xlim = c(1.1,24.9), ylim = c(1.1,(n_var-0.1))) + 
  xlab("Factors") + 
  ylab("Metabolites") + 
  ggtitle("Cluster Metabolites") + 
  theme(axis.title.x = element_text(size=8), 
        axis.title.y = element_text(size=8),
        axis.text.x = element_text(size=6),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust = 0.5))
p

# Create a table with significant metabolites only
tmp_table = data.table(colnames(X),
                       apply(intdl$mainEffectSamps,1,mean),
                       t(apply(intdl$mainEffectSamps,1,quantile,probs = c(0.025,0.975))))
tmp_table[,sig := ifelse(sign(`2.5%`)==sign(`97.5%`),1,0)]
tmp_table[sig == 1 & V1 != "age_birth"]

tmp = fread(file = "J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/1_metabolite-details.csv")
table_sig = merge(tmp,tmp_table[,.(V1,Mean = V2,`2.5%`,`97.5%`,sig)],
                  by.x = "comp_id", by.y = "V1")

datatable(table_sig, filter = 'top', options = list(
  pageLength = 5, autoWidth = TRUE,lengthChange = FALSE))

# Plot of the significant metabolites usign FIN
ggplot(tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)], aes(V1,V2)) + 
  geom_hline(yintercept=0, lty=2, lwd=1, colour="grey50") +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), 
                lwd=0.5, width=0) +
  geom_point(size=2) +
  scale_x_discrete(limits = tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)][,V1]) + 
  coord_flip() +
  guides(colour=FALSE) +
  ggtitle("Cluster Metabolites") + 
  labs(x="Coefficient", y="Value") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.grid = element_line(color = "grey"),
        axis.text.x = element_text(size=6),
        axis.text.y = element_text(size=6))

# Quadratic term 
inter_m_mat = apply(intdl$omegaSamps, c(1,2), mean)
inter_q1_mat = apply(intdl$omegaSamps, c(1,2), quantile, c(0.025))
inter_q2_mat = apply(intdl$omegaSamps, c(1,2), quantile, c(0.975))

inter_m_mat = as.data.table(inter_m_mat)
inter_m_mat[,names := colnames(X)]
setnames(inter_m_mat,names(inter_m_mat),c(colnames(X),"names"))
inter_m_mat = melt(inter_m_mat)

inter_q1_mat = as.data.table(inter_q1_mat)
inter_q1_mat[,names := colnames(X)]
setnames(inter_q1_mat,names(inter_q1_mat),c(colnames(X),"names"))
inter_q1_mat = melt(inter_q1_mat)

inter_q2_mat = as.data.table(inter_q2_mat)
inter_q2_mat[,names := colnames(X)]
setnames(inter_q2_mat,names(inter_q2_mat),c(colnames(X),"names"))
inter_q2_mat = melt(inter_q2_mat)

inter_m_mat = as.data.table(inter_m_mat)
inter_q1_mat = as.data.table(inter_q1_mat)
inter_q2_mat = as.data.table(inter_q2_mat)

data_inter = Reduce(function(x, y) merge(x, y, 
    by=c("names", "variable")), list(inter_q1_mat[,.(names, variable=as.character(variable), value)],
                                     inter_m_mat[,.(names, variable=as.character(variable), value)],
                                     inter_q2_mat[,.(names, variable=as.character(variable), value)]))
data_inter[,check := ifelse(sign(value.x) == sign(value),1,0)]

data_inter_tmp = data_inter[names %in% metabolites_list & variable %in% metabolites_list]
data_inter_tmp[, value_plot := ifelse(check == 0, NA, value.y)]
data_inter_tmp = data_inter_tmp[names %in% tmp_table[sig==1 & V1 != "age_birth",V1] & 
                                variable %in% tmp_table[sig==1  & V1 != "age_birth",V1]]
data_inter_tmp[, variable := factor(x = variable,
                                    labels = tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)][,V1])]
data_inter_tmp[, names := factor(x = names,
                                    labels = tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)][,V1])]
setkey(data_inter_tmp,names,variable)

tmp_k = matrix(NA,46,46)
tmp_k[1,] = c(rep(1,1),rep(0,45))
for(i in 2:46){
  tmp_k[i,] = c(rep(1,i),rep(0,46-i))
}
kkk = as.vector(t(tmp_k))
data_inter_tmp[,tri_plot := kkk]
data_inter_tmp[, value_plot := ifelse(tri_plot==1,value_plot,NA)]

data_inter_tmp[,value_plot := cut(data_inter_tmp$value_plot,c(min(data_inter_tmp$value_plot,na.rm = T),
                                                              0,5,10,
                                                              max(data_inter_tmp$value_plot,na.rm = T)))]

ggplot(data_inter_tmp, aes(names, variable, fill= value_plot)) + 
  geom_tile() +
  scale_y_discrete(limits = unique(data_inter_tmp$names)) +
  scale_x_discrete(limits = rev(unique(data_inter_tmp$names))) +
  scale_fill_manual(values=c("#f4d219","#99CCFF","#6699CC","#336699")) +
  ggtitle("Interaction Metabolites") +
  labs(x="", y="") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size=6,angle = 90,hjust=0.95,vjust=0.2),
        axis.text.y = element_text(size=6))

# Interactions metabolites-covariates
data_cov_tmp = data_inter
data_cov_tmp[, value_plot := ifelse(check == 0, NA, value.y)]
data_cov_tmp = data_cov_tmp[names %in% tmp_table[sig==1 & V1 != "age_birth",V1] & 
                                variable %in% c(covariates[c(5,7,9,12)])]
data_cov_tmp[, names := factor(x = names,
                               labels = tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)][,V1])]
data_cov_tmp[, variable := factor(x = variable,
                               labels = c(covariates[c(5,7,9,12)]))]
setkey(data_cov_tmp,names)

data_cov_tmp[,value_plot := cut(data_cov_tmp$value_plot,c(min(data_cov_tmp$value_plot,na.rm = T),-10,-5,0,
                                                          5,10,
                                                          max(data_cov_tmp$value_plot,na.rm = T)))]

ggplot(data_cov_tmp, aes(y = names, x = variable, fill= value_plot)) + 
  geom_tile() +
  scale_fill_simpsons() + 
  ggtitle("Interaction Metabolites-Covariates") +
  labs(x="", y="") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=6))
```

```{r plot,echo=F, warning=FALSE, message=FALSE, fig.align='center'}
dt_boxplot = melt(DT[,as.character(unique(data_cov_tmp$names)),with=F])

ggplot(dt_boxplot, aes(x=variable, y=value)) + 
  geom_boxplot(fill = "grey70",color = "black",alpha = 0.5) + 
  scale_x_discrete(limits = unique(data_inter_tmp$names)) + 
  coord_flip() +
  labs(title = "Value of significant Metabolites") +
  theme_bw() + 
  xlab("") +
  ylab("") +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7))

cat("Pearson correlation matrix ")
corrplot(cor(met[,tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = F)][,V1],with=F]),
                tl.col = "black", tl.cex = .5, tl.pos = "lt")
cat("Spearman correlation matrix ")
corrplot(cor(met[,tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = F)][,V1],with=F], 
                   method = "spearman"),
                tl.col = "black", tl.cex = .5, tl.pos = "lt")

# Create distance matrx
met_sim_dt = as.matrix(t(DT[,table_sig[sig==1,comp_id],with=F]))
corMat = negDistMat(met_sim_dt,r=2)
D = corMat/min(corMat)
spe.o = order.single(1-D)

col3 <- colorRampPalette(c("red","yellow","hotpink","white", "cyan"))
col3 <- colorRampPalette(c("black","black","white"))
corrplot(D[spe.o,spe.o], method = "color",is.corr = F, col = col3(40), 
         tl.col = "black", tl.cex = 0.5)	

```

# MWAS Analysis

Here we run a single metablites analysis and we select the metabolites that are significant in both of the analysis

```{r single, echo=F, warning=FALSE, message=FALSE, fig.align='center'}
library(BAS)
tmp = lapply(1:length(metabolites_list),function(i){
  lm_formula = as.formula(paste0("telomere_cordblood ~ ",metabolites_list[i]," + gestage_serum + age_birth + white + black + other + hispanic + educ + mbmi_prepreg"))
  met_bas <- bas.lm(lm_formula,
                  data = data.frame(telomere_cordblood = y,X),
                  prior = "BIC",
                  modelprior = uniform())
  lm_model = lm(lm_formula, data = data.frame(telomere_cordblood = y,X))
  lm_table = data.frame(metabolite = metabolites_list[i],
                        t(confint(coefficients(met_bas))[2,]),
                        t(coefficients(summary(lm_model))[2,]),
                        confint(lm_model, level = 0.95)[2,1],
                        confint(lm_model, level = 0.95)[2,2])
})
tmp = rbindlist(tmp)
tmp = tmp[,.(metabolite, B_estimates = beta, cri025 = X2.5., cri975 = X97.5., F_estimates = Estimate,
             SE = `Std..Error`,`P-value` = `Pr...t..`,ci025 = confint.lm_model..level...0.95..2..1.,
             ci975 = confint.lm_model..level...0.95..2..2.)]
tmp[, bonf := p.adjust(tmp$`P-value`, method = "bonferroni")]
tmp[, fdr := p.adjust(tmp$`P-value`, method = "fdr")]

tmp = readRDS("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Results/Prepreg_bmi/BW_25_scaled_mwas_all_results_bmi.rds")

p = ggplot(tmp, aes(F_estimates, -log10(`P-value`))) +
  geom_point(alpha=0.5, cex = 2, color = "grey50") +
  geom_hline(yintercept = -log10(0.00018), linetype = 2) +
  geom_hline(yintercept = -log10(0.05), linetype = 2) +
  theme_bw() 
k = p + 
  xlab("beta") + ylab("-log10(pvalue)") +
  ggtitle(paste0("")) +
  labs(title = "Volcano plot metabolites",
       subtitle = (" correction"),
       caption = "Data source: Metabolomics") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size=15),
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size = 10)) 
print(k)

# CI single analysis metabolites 
tmp_met_ci = tmp[metabolite %in% as.character(table_sig[sig==1,comp_id])]
tmp_met_ci[, metabolite := factor(x = metabolite, levels = tmp_table[sig == 1 & V1 != "age_birth"][order(V2, decreasing = T)][,V1])]
ggplot(tmp_met_ci, aes(x=F_estimates, y=metabolite)) + 
  geom_point(size = 1) + 
  geom_pointrange(aes(xmin=ci025, xmax=ci975), size = 0.5 ) +
  geom_vline(xintercept = 0, lty=2, lwd=0.5, color = "grey50") + 
  labs(title = "Metabolites Regression",
       subtitle = "Regression one by one",
       caption = "Data source: Metabolomics") +
  xlab("Coefficients") + 
  ylab("Metabolites") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size=15),
        axis.text.x = element_text(size=6),
        axis.text.y = element_text(size = 6),
        panel.grid = element_line(color = "grey"))  
```

```{r qqplot_pvalues, echo=F, warning=FALSE, message=FALSE, fig.align='center'}
qqplot(runif(1110),tmp$`P-value`, ylab = "Observed p-value",xlab = "Uniform p-value",
       main = "Genomic inflation factor")
abline(0,1,lty = 2)
ks.test(tmp$`P-value`,'punif')

#Lambda = genomic inflation factor
pvalue <- tmp$`P-value`
chisq <- qchisq(1 - pvalue, 1)
lambda <- median(chisq) / qchisq(0.5, 1)

# Confidence Interval
SE.median <- qchisq(0.975, 1) * (1.253 * ( sd(chisq) / sqrt( length(chisq) ) ) )
paste0("Lambda inflation factor: ",round(lambda,3)," (",
       round(lambda - (SE.median / qchisq(0.5, 1)),3),"; ",
       round(lambda + (SE.median / qchisq(0.5, 1)),3),")")



```

```{r scatter, echo=F, fig.width=12, fig.height=12}
hormones = read_excel("J:/PM/Colicino_Lab/Manuscripts/2020_Metabolomics_Ferrari/Data/hormones.xlsx")
hormones = data.table(hormones)
hormones = hormones[complete.cases(hormones)]
DT_scatter = DT_unscaled[,c("gestage",hormones$`Coumpound ID`),with=F]

plt = vector("list",16)
for(i in 1:NROW(hormones)){
  DT_scatter_plot = DT_scatter[,.(gestage,value = get(hormones$`Coumpound ID`[i]))]
  plt[[i]] = ggplot(DT_scatter_plot, aes(x = gestage, 
                                    y = value)) + 
    geom_point() + geom_smooth() +
    theme_bw() + 
    labs(title = paste0(hormones$`Coumpound ID`[i]," - ",
                        hormones$`Super Pathway`[i]),
         subtitle = paste0(hormones$Biochemical[i])) +
    xlab("Gestational Age") + ylab(paste0(hormones$`Coumpound ID`[i])) +
    theme(plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5,size = 8))
}
grid.arrange(plt[[1]],plt[[2]],plt[[3]],plt[[4]],
             plt[[5]],plt[[6]],plt[[7]],plt[[8]],
             plt[[9]],plt[[10]],plt[[11]],plt[[12]],
             plt[[13]],plt[[14]],plt[[15]],plt[[16]],
             ncol = 4, nrow = 4)



```




